---
title: "Unemployment Analysis by County"
subtitle: "Kyle Kirk's Data Science Portfolio" #replace blank with your name
output:
  html_document:
    df_print: paged
---

## Abstract

After looking over a few datasets from Nate Silver, data.gov, bigrquery and google. After exhausting many avenues of data I found a solid lead on the USDA website : (USDA.)[https://www.ers.usda.gov/data-products/county-level-data-sets/download-data/]

Using this data I hope to find a deeper parallel between unemployment and the increase in automation technology.

***
### 1. Discovery and Data Preparation
***

The first thing to be done is to load the libraries that we will be using during our investigation of the data.

```{r}

```

The second thing is to import the data from the USDA government website.

```{r}
income_ineq <- read_excel("~/Downloads/Unemployment.xls")
```
Provide a one-sentence introduction with what is covered by the first deliverable along with a link to another html page that gives the details.

### 2. Model Planning and Building

Coming soon...

### 3. Results and Operationalization

Coming soon...

